Topological Equivariant Artist Model

### 2021-03-18 paper/paper.tex
equivariant monoid figure

 \subsubsection{Sheafs: Gluing Sections }
 \label{sec:math:data:sheaf}
 The restriction maps of a sheaf describe how local sections $\iota^{*}\dsection$ over a neighborhood $U$ around \dbasepoint\ can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}. This puts a continuous structure on local sections, which allows for defining a section over a subset in \dbase. The inclusion map $\iota: U \rightarrow \dbase$ pulls \dtotal\ over $U$ 
 \begin{equation}
     \label{eq:sheaf}
     \begin{tikzcd}
         \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
         U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
     \end{tikzcd}
 \end{equation}
 such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$. Gluing together $\iota^*\dsection$ sections is necessary for navigation techniques such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006} and dynamically updated visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

@@

\subsubsection{Visual Component Maps}
\label{sec:math:artist:nu}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\columnwidth]{equivariance_nu.png}
  \caption{In this artist, \vchannel\ maps the strings to the emojis. This \vchannel\ is equivariant because the monoid actions, represented by the colored arrows, are the same on the input \dsection\ and output \vsection\ sets.}
  \label{fig:math:artist:nu}
\end{figure}
 We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. 
A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes such that 
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$. Therefore, without a loss of generality, we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function.

The \vchannel\ illustrated in \autoref{fig:math:artist:nu} is an equivariant mapping from \textbf{Strings} to emojis, as seen in the identical actions, represented as arrows, on the data and emojis. More generally, for the Steven's scales\cite{stevensTheoryScalesMeasurement1946} an equivariant \vchannel\ has the following properties
\begin{table}[H]
  \begin{tabulary}{\columnwidth}{llL}
      nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
      ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
      interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
      ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\
  \end{tabulary}
\end{table}
\noindent where the middle column is the group structure. Therefore, a \vchannel\ is equivariant if it satisfies the listed condition. For example, given a transform $\vchannel_{i}(x) = .5$ on interval data, it must commute under translation. Testing this constraint with a translation action $t(x) = x+2$
\begin{align*}
  \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + 2\\
  .5 &\neq .5 + 2
\end{align*}
\vchannel\ does not commute and is therefore invalid. The constraints on \vchannel\ can be embedded into the artist such that the \vchannel\ functions can test for equivariance and also provide guidance on constructing new \vchannel\ functions.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{base_q}
  \caption{The coordinates $(\gx, \gy)$ dictate the color of the region in prerender space \gbase\, which over the whole disk generates the graphical mark. The line fiber is thickened with the derivative because the tangent the line needs to be pushed perpendicular to the tangent of (xpos, ypos) in order to have visible thickness.}
  \label{fig:math:artist:base}
\end{figure}

>>>>

 \subsubsection{Sheafs}
 \label{sec:math:data:sheaf}
 The restriction maps of a sheaf describe how local sections $\iota^{*}\dsection$ over a neighborhood $U$ around \dbasepoint\ can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}. This puts a continuous structure on local sections, which allows for defining a section over a subset in \dbase. The inclusion map $\iota: U \rightarrow \dbase$ pulls \dtotal\ over $U$ 
 \begin{equation}
     \label{eq:sheaf}
     \begin{tikzcd}
         \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
         U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
     \end{tikzcd}
 \end{equation}
 such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$. Gluing together $\iota^*\dsection$ sections is necessary for navigation techniques such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006} and dynamically updated visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

@@

\subsubsection{Visual Component Maps}
\label{sec:math:artist:nu}
 We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. 
A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes such that 
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$. Therefore, without a loss of generality, we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function.

\begin{figure}[htb]
  \centering
  \includegraphics[width=1\columnwidth]{monoid_equivariant.png}
  \caption{fill in this caption tomorrow}
  \label{fig:math:artist:nu}
\end{figure}

\autoref{fig:math:artist:nu} illustrates a \vchannel\ that preserves partial order relations because 
ou would have to state that the monoid of monitonic actions is equivalent to the partial order realationship
ok
so that the order is preserved IF and only IF it is an equivariant map.
I think that is true but it would be nice to have a reference.
but then you don't have to really plot out m.

Generally, for the Steven's scales\cite{stevensTheoryScalesMeasurement1946} an equivariant \vchannel\ has the following properties
\begin{table}[H]
  \begin{tabulary}{\columnwidth}{llL}
      nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
      ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
      interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
      ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\
  \end{tabulary}
\end{table}
\noindent where the middle column is the group structure. Therefore, a \vchannel\ is equivariant if it satisfies the listed condition.

@@

\begin{figure}
  \centering
  \includegraphics[width=.5\columnwidth]{base_q}
  \caption{The coordinates $(\gx, \gy)$ dictate the color of the region in prerender space \gbase\, which over the whole disk generates the graphical mark. The line fiber is thickened with the derivative because the tangent the line needs to be pushed perpendicular to the tangent of (xpos, ypos) in order to have visible thickness.}
  \label{fig:math:artist:base}
\end{figure}

### 2021-03-18 slides/main.tex
equivariant monoid figure

\begin{frame}{Variable types are dimensions of the fiber}
    \begin{figure}[H]
        \begin{subfigure}{.49\textwidth}
            \includegraphics[width=\textwidth]{figures/math/temp_1k.png}
            \caption{}
            \label{fig:fiber_example_plane}
        \end{subfigure}
        \begin{subfigure}{.49\textwidth}
            \includegraphics[width=\textwidth]{figures/math/temp_3f.png}
            \caption{}
            \label{fig:fiber_example_cube}
        \end{subfigure}
        \caption{}
        \label{fig:data_fiber_example}
    \end{figure}
    \begin{description}
        \item[\ref{fig:fiber_example_plane}]  $\dfiber=\reals\times\reals$, (time, temperature)
        \item[\ref{fig:fiber_example_cube}]  $\reals\times\realsp\times\reals$, (time, wind=(speed, direction))
    \end{description}
\end{frame}

@@

\begin{frame}{Data connectivity is encoded as the base space}
    \begin{figure}[H]
        \begin{subfigure}{.49\textwidth}
            \includegraphics[width=\textwidth]{figures/math/temp_1k.png}
            \caption{}
            \label{fig:base_example_discrete}
        \end{subfigure}
        \begin{subfigure}{.49\textwidth}
            \includegraphics[width=\textwidth]{figures/math/temp_2f.png}
            \caption{}
            \label{fig:base_example_continuous}
        \end{subfigure}
        \caption{}
        \label{fig:base_example}
    \end{figure}
    \begin{description}
        \item[\ref{fig:base_example_discrete}] data is discrete 0D points
        \item[\ref{fig:base_example_continuous}]  data is lies on the 1D continuous interval \dbase\
    \end{description}
\end{frame}

@@

\begin{frame}{Example: Nominal Equivariance}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/equivariance_nu.png}
        \caption{The actions on the text data are the same as the actions on the visual representation of that data as emojis.}
    \end{figure}
\end{frame}
\begin{frame}{Measurement Scale Groups}

>>>>

\begin{frame}{Variable types are dimensions of the fiber}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/fiber.png}
        \caption{}
        \label{fig:data_fiber_example}
    \end{figure}
    \begin{description}
        \item[plane]  $\dfiber=\reals\times\reals$, (time, temperature)
        \item[cube]  $\reals\times\realsp\times\reals$, (time, wind=(speed, direction))
    \end{description}
\end{frame}

@@

\begin{frame}{Data connectivity is encoded as the base space}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
        \caption{}
        \label{fig:base_example}
    \end{figure}
    \begin{description}
        \item[points] data is 0D discrete
        \item[line]  data is lies on the 1D continuous interval \dbase\
    \end{description}
\end{frame}

@@

\begin{frame}{Example: Nominal Equivariance}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}
\begin{frame}{Measurement Scale Groups}
